Elimination Of Jump Objects

### 2022-06-18 paper/skeleton.tex
Update skeleton.tex

\documentclass[sigplan,review,11pt,nonacm,natbib=false]{acmart}
\settopmatter{printfolios=false,printccs=false,printacmref=false}
\usepackage[maxnames=1,minnames=1,maxbibnames=100,natbib=true,citestyle=authoryear,bibstyle=authoryear,doi=false,url=false,isbn=false,isbn=false]{biblatex}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{ffcode}

@@

% Keywords command
\providecommand{\keywords}[1]
{
  \small
  \textbf{\textit{Keywords:}} #1
}

\definecolor{codegray}{rgb}{0.5,0.5,0.5}

@@

Some objects in EO programs may need to be platform specific and can’t be composed from other existing objects—they are called atoms. 
So all jump statements are implemented as a
single atom-object GOTO. The implementation of goto-object in EO implies two types of jumps - goto forward and goto backward. So we should consider transformations for both cases.

@@

\emph{Definition:} The \emph{nesting} of the current object means the number of objects that contain the current object directly, or that contain objects that contain the current object.

@@

@@

$\sigma$ - Means that there can be an arbitrary number of objects, a chain of nested objects or there can be no objects at all.

$\delta$ - Means that there should necessarily be an object that returns a non-NULL value when accessed.

$\mu$ - Means that there must be an object that returns a boolean value (TRUE/FALSE) when executed.

@@

@@

goto
  [g]
    |\sgm{1}|
      g.forward |\dlt{1}|
      stdout "Will never be printed"
\end{ffcode}

@@

\subsubsection{Simple Forward}
This is the simplest example of using goto-forward:

\begin{ffcode}
goto
  [g]
    g.forward |\dlt{1}|
\end{ffcode}

Note that in this example, g.forward is simply an attribute of the abstract object g and is executed as soon as the queue reaches it.
Our transformation only considers the jumps that are made when a condition for it is fulfilled. In cases where there is unconditional jump, first of all we transform it into conditional jump by adding an additional if-statement - if (TRUE).
This will not violate the program logic, because without the added if-statement, the jump would always be executed, and statements coming after this jump with a nesting not less that the nesting of the abstract object \emph{g} would never be executed:

\begin{ffcode}
goto
  [g]
    if.
      TRUE
      g.forward |\dlt{1}|
      TRUE
\end{ffcode}

\subsubsection{Goto forward wrapped in an object}
So, after the conversion described in section 3.1.1, in general cases, forward jumps will look like this:
\begin{ffcode}
goto
  [g]
    |\sgm{1}| |$\label{ln:ret0}$|
      |\sgm{2}|
      if.
        |\m{1}|
        g.forward |\dlt{1}|
        |\dlt{2}|
      |\sgm{3}| |$\label{ln:ret1}$|
\end{ffcode}

@@

It is not difficult to notice that statements which goto-forward "jumps through" will be executed if the condition in the if-statement is not equal to TRUE.
Otherwise statements following the goto object itself are executed.
So we can swap statements and invert the condition into if-statement to get rid of the \emph{g.forward} attribute without violating program logic (Look at Image1).

Also, after replacing the jump, we must create an additional \emph{flag} object, which will change if the jump condition is met, i.e. in the new if-statement, the 'ELSE' branch will be executed. We need this so that when a series of nested objects are nested one in another, when the jump condition is met, subsequent objects whose nesting is less than the nesting of the original object containing goto forward and more than the nesting of an abstract object \emph{g} and following after the original object are not executed.

By swapping the statements that follow the if-statement body and after the jump itself, as described above, we get the following replacement for this example: the \m{1} object must be inverted with not. (\m{1}); in 'THEN' brunch of if-statement creates an object \emph{seq} with \dlt{2} and \sgm{3} inside of it; the 'ELSE' branch an object \emph{seq} with \dlt{1} object that was previously called when the jump is made and changes the value of the \emph{flag} object to 1.

{INSERT AN IMAGE1}

Thus, after transformation, the program will look like this:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{1}||$\label{ln:ret2}$|
  |\sgm{2}|
  if.
    |\m{1}|.not
    |\dlt{2}|
    seq
      |\dlt{1}|
      flag.write 1
  if.
    (eq. (flag 1)).not
    |\sgm{3}|
    TRUE|$\label{ln:ret3}$|
\end{ffcode}

@@

Now let's look at an example program with nested objects containing goto forward:
\begin{ffcode}
goto
  [g]
    |\sgm{0}|
      X
      |\sgm{4}|
\end{ffcode}

@@

A goto-object can also be called inside such a while-loop. In this case, the transformation is slightly different from that described in Section 3.1.2, because the behavior of while is different from other objects in the language.

In general, using goto forward inside a while-loop will look like this:
\begin{ffcode}
goto
  [g]
    |\sgm{0}||$\label{ln:ret4}$|
      while.
        |\m{0}|
        |\sgm{1}|
          |\sgm{2}|
          if.
            |\m{1}|
            g.forward |\dlt{1}|
            |\dlt{2}|
          |\sgm{3}||$\label{ln:ret5}$|
\end{ffcode}

@@

Thus, transformation will look like this:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{0}||$\label{ln:ret6}$|
  while.
    and.
      (eq. (flag 1)).not
      |\m{0}|
    |\sgm{1}|
      |\sgm{2}|
      if.
        |\m{1}|.not
        |\dlt{2}|
        seq
          |\dlt{1}|
          flag.write 1
      if.
        (eq. (flag 1)).not
        |\sgm{3}|
        TRUE|$\label{ln:ret7}$|
\end{ffcode}

@@

It is worth noting that if goto forward represent several nested while-loops, then in each of them should be added to the condition of the loop check whether the flag is equal to 1, as in the inner while, from where the jump is made. So, every while that has a nesting less than the original nesting and more than the nesting of the abstract object \emph{g} will end without performing the next iteration.

For example, a program with nested while-loops may look like this:
\begin{ffcode}
goto
  [g]
    while.
      |\m{9}|
      |\sgm{9}|
        Y
        |\sgm{4}|
\end{ffcode}

@@

goto
  [g]
    |\sgm{1}|
      stdout "Will be printed forever"
      g.backward
\end{ffcode}

Same as with goto forward, there is the goto atom-object, the abstract object \emph{g}, and the g.backward attribute. Several abstract objects such as \emph{g} can be declared inside of goto-object, and the attribute \emph{.backward} can be called to each of them.

@@

\begin{ffcode}
goto
  [g]
    g.backward
\end{ffcode}

This is identical to that discussed in Section 3.1.1. In cases where the jump is performed without condition, you need to add an additional if-statement, similar to the example in Section 3.1.1 but with goto-backward:

\begin{ffcode}
goto
  [g]
    if.
      TRUE
      g.backward
      TRUE
\end{ffcode}

\subsubsection{Goto backward wrapped in an object}
After the conversion decribed in Section 3.2.1, in general cases, backward jumps will look like this:

@@

Note that goto-backward works the same as do-while construction in other languages: the first iteration of the new loop is always executed, and the next only if the condition is TRUE.
So, all objects that occurs between the "jump" itself and the declaration of an abstract object \emph{g} should be executed before the new while loop once, and all subsequent times were executed depending on the condition of the loop (Look at Image2).

As soon as the execution of the program reaches the jump condition, the first iteration of the new do-while loop will already be executed.
After that, if the jump conditions are met, all objects that are between the declaration of the abstract object \emph{g} and the jump itself will be executed again. And that means that we can create a while loop at the point of the jump, so that every time we check the condition, we don't make a jump, but just start a new iteration of the loop.

Inside the body of the new while loop, we copy all objects that are inside the abstract object g - with nesting at least the object g itself, except for the call of the jump itself with its condition.
At the point of the jump, we should create an if-statement with an inverted jump condition: in the 'THEN' branch there will be objects that were returned in the 'ELSE' branch of the original if for the jump; in the 'THEN' branch, the value of the flag object will be changed, which will mean that the condition for the jump is met, and all subsequent objects with nesting at least the nesting of the added while loop will not be executed.
To do this, for all these objects, going down the nesting from the object where the jump was made to the while loop, we need to add an if-statement that will check whether the condition for the jump has been met before.

This transformation does not violate the program logic, because if the jump condition is fulfilled, then all the objects going after the jump itself will not be executed due to the added if-statements.

{INSERT AN IMAGE2}

So, after transformation, the program will look like this:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{1}||$\label{ln:ret10}$|
  |\sgm{2}|
  if.
    |\m{1}|.not
    |\dlt{1}|
    while.
      |\m{1}|
      seq
        |\sgm{1}|
          |\sgm{2}|
          if.
            |\m{1}|.not
            |\dlt{1}|
            flag.write 1
          if.
            (eq. (flag 1)).not
            |\sgm{3}|
            TRUE
        flag.write 0
  |\sgm{3}||$\label{ln:ret11}$|

@@

\begin{ffcode}
goto
  [g]
    |\sgm{0}|
      W
      |\sgm{4}|
\end{ffcode}

Applying transformation, we get:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{0}|
  W'
  if.
    (eq. (flag 1)).not
    |\sgm{4}|
    TRUE
\end{ffcode}

\subsubsection{Goto backward inside the while-loop}
Goto backward, like goto forward, can be called inside a while-loop body.
As discussed in section 3.1.3, the behavior of while is different from other objects in the language. Therefore, the transformation will be slightly different from that described in section 3.2.2.

In general cases, a goto backward call inside a while loop will look like this:

\begin{ffcode}
goto
  [g]
    |\sgm{0}||$\label{ln:ret12}$|
      while.
        |\m{0}|
        |\sgm{1}|
          |\sgm{2}|
          if.
            |\m{1}|
            g.backward
            |\dlt{1}|
          |\sgm{3}||$\label{ln:ret13}$|

@@

\begin{ffcode}
memory > flag
flag.write 0
|\sgm{0}||$\label{ln:ret14}$|
  while.
    |\m{0}|
    |\sgm{1}|
      |\sgm{2}|
      while.
        |\m{1}|
        |\sgm{0}|
          while.
            and.
              (eq. (flag 1)).not
              |\m{0}|
            |\sgm{1}|
              |\sgm{2}|
              if.
                |\m{1}|.not
                |\dlt{1}|
                flag.write 1
              if.
                (eq. (flag 1)).not
                |\sgm{3}|
                TRUE
      |\sgm{3}||$\label{ln:ret15}$|

@@

Let's take a look on a program with nested while-loops:
\begin{ffcode}
goto
  [g]
    while.
      |\m{9}|
      |\sgm{9}|
        Z
        |\sgm{4}|
\end{ffcode}

After transformation it will look like:
\begin{ffcode}
memory > flag
flag.write 0
while.
  |\m{9}|
  |\sgm{9}|
    Z'
    if.
      (eq. (flag 1)).not
      |\sgm{4}|
      TRUE
\end{ffcode}

>>>>

\documentclass[sigplan,review,11pt,nonacm,natbib=false]{acmart}
\settopmatter{printfolios=false,printccs=false,printacmref=false}
\usepackage[maxnames=1,minnames=1,maxbibnames=100,natbib=true,citestyle=authoryear,bibstyle=authoryear,doi=false,url=false,isbn=false,isbn=false]{biblatex}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{ffcode}
\usepackage{amsmath}

\newtheorem{definition}{Definition}[section]

@@

@@

Some objects in EO programs may need to be platform specific and can’t be composed from other existing objects—they are called \emph{atoms}. 
So all jump statements are implemented as a
single \emph{atom}-object GOTO. The implementation of goto-object in EO implies two types of jumps - goto forward and goto backward. So we should consider transformations for both cases.

@@

\begin{definition} The \emph{nesting} of the current object means the number of objects that contain the current object directly, or that contain objects that contain the current object.
\end{definition}

@@

\vspace{5mm}

@@

\begin{itemize}
\item $\sigma$ - Denotes either an object, a chain of nested objects or the absence of an object.

\item $\delta$ - Denotes a not-empty object or a chain of nested objects that return a not-NULL value when accessed.

\item $\mu$ - Denotes an object or a chain of nested objects that return a boolean value (TRUE/FALSE) when executed.
\end{itemize}

@@

We will also look at our transformations in the form of $\varphi$-calculus, on which EO is based. The proposed $\varphi$-calculus represents an object model through data and objects, while operations with them are possible through abstraction, application, and decoration.
The calculus introduces a formal apparatus for manipulations with objects.

$\varphi$-calculus uses an alternative “arrow notation” to denote an object in a more compact way, where free attributes of an abstract object stay in the parentheses on the left side of the mapping symbol $\mapsto$ and pairs, which represent bound attributes, stay on the right side, in double-square brackets.

@@

goto
  [g]
    |\sgm{1}| > @
      g.forward |\dlt{1}|
      stdout "Will never be printed"
\end{ffcode}

@@

\subsubsection{Simple Forward}
Consider the simplest example of using goto-forward:

\begin{ffcode}
goto
  [g]
    g.forward |\dlt{1}| > @
\end{ffcode}

In terms of $\varphi$-calculus, this code will appear as:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \varphi \mapsto \ff{g}.\ff{forward} (\delta_1) \rrbracket ) \\
\end{split}
\end{equation}

Note that in this example, g.forward is simply an attribute of the abstract object \emph{g} and is executed as soon as the queue reaches it.
Our transformation only considers the jumps that are made when a condition for it is fulfilled. In cases where there is unconditional jump, first of all we transform it into conditional jump by adding an additional if-statement - if (TRUE).
This will not violate the program logic, because without the added if-statement, the jump would always be executed, and statements coming after this jump with a nesting not less that the nesting of the abstract object \emph{g} would never be executed:

\begin{equation}
\begin{split}
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.( TRUE, \ff{g}.\ff{forward}(\delta_1), TRUE ) \rrbracket) 
\end{split}
\end{equation}

Same example on EO:

\begin{ffcode}
goto
  [g]
    if. > @
      TRUE
      g.forward |\dlt{1}|
      TRUE
\end{ffcode}

\subsubsection{Goto forward wrapped in an object}

After the conversion described in section 3.1.1, the simplest call of forward jump will be presented as:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.(\mu_1, \ff{g}.\ff{forward}(\delta_1), \delta_2) \rrbracket) \\
\end{split}
\end{equation}

The transformation under consideration involves the following steps:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.(\mu_1.\ff{not}, \ff{g}.\ff{forward}(\delta_1), \delta_2) \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.(\mu_1.\ff{not}, \delta_2, \delta_1) \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{if.}(\mu_1.not, \delta_2, \delta_1) \\
\end{split}
\end{equation}

\begin{theorem}
Expression 3 is semantically equivalent to expression 6:
\end{theorem}

\begin{equation} \nonumber
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.(\mu_1, \ff{g}.\ff{forward}(\delta_1), \delta_2) \rrbracket) \\
\end{split}
\end{equation}

\hline

\begin{equation} \nonumber
\begin{split}
& \ff{if.}(\mu_1.not, \delta_2, \delta_1) \\
\end{split}
\end{equation}

To prove that the transformation does not violate program logic, let's consider it with the following abstract case of calling goto forward.

Consider the following code:

\begin{ffcode}
goto
  [g]
    |\sgm{1}| > @ |$\label{ln:ret0}$|
      |\sgm{2}|
      if.
        |\m{1}|
        g.forward |\dlt{1}|
        |\dlt{2}|
      |\sgm{3}| |$\label{ln:ret1}$|
\end{ffcode}

@@

It is not difficult to notice that statements which goto-forward "jumps through" will be executed if the condition in the if-statement is not equal to TRUE.
Otherwise statements following the goto-object itself are executed.
So we can swap statements and invert the condition into if-statement to get rid of the \emph{g.forward} attribute without violating program logic.

Also, after replacing the jump, we must create an additional \emph{flag} object, which will be changed if the jump condition is met, i.e. when in the new if-statement, the 'ELSE' branch will be executed - value of flag will change. This is required so that all objects following the if-statement that are attributes of the abstract object g are not executed.
So, by wrapping each object following the jump condition and whose nesting is at least that of the g object in an if-statement with a flag value check condition, we make sure that these objects will not be executed when the jump condition is met.

By swapping the statements that follow the if-statement body and after the jump itself, as described above, we get the following replacement for this example: the \m{1} object must be inverted with (\m{1}).not; in 'THEN' brunch of if-statement places \dlt{2}; in the 'ELSE' brunch creates new object seq, consisting of two elements - \dlt{1} and flag's value change, because the jump condition is satisfied.

Thus, after transformation, the program converts to:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{1}||$\label{ln:ret2}$|
  |\sgm{2}|
  if.
    |\m{1}|.not
    |\dlt{2}|
    seq
      |\dlt{1}|
      flag.write 1
  if.
    (eq. (flag 1)).not
    |\sgm{3}|
    TRUE|$\label{ln:ret3}$|
\end{ffcode}

@@

Now let's look at an example program with nested objects containing goto forward:
\begin{ffcode}
goto
  [g]
    |\sgm{0}| > @
      X
      |\sgm{4}|
\end{ffcode}

@@

The simplest way to call goto forward inside a while-loop is represented as:

\begin{equation}
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\mu_0, \\
& \quad \quad \ff{if}.(\mu_1, \ff{g}.\ff{forward}(\delta_1), \delta_2) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

The transformation under consideration involves the following steps:

%%\begin{equation}
%%\begin{split}
%%& \ff{memory}(flag), \\
%%& flag.\ff{write}(0), \\
%%& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} %%\mapsto \varnothing, \\
%%& \quad \varphi \mapsto \ff{while}.(\mu_0, \\
%%& \quad \ff{if}.(\mu_1, %%\ff{g}.\ff{forward}(\delta_1), \delta_2) \\
%%& \quad )\rrbracket) \\
%%\end{split}
%%\end{equation}

\begin{equation}
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\ff{and}.( \\
& \quad \quad \mu_0, \ff{eq}.(flag  1).\ff{not}), \\
& \quad \quad \ff{if}.(\mu_1,  \ff{g}.\ff{forward}(\delta_1), \delta_2) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\ff{and}.( \\
& \quad \quad \mu_0, \ff{eq}.(flag  1).\ff{not}), \\
& \quad \quad \ff{if}.(\mu_1.\ff{not}, \delta_2, \\
& \quad \quad \quad \ff{seq}( \delta_1, flag.\ff{write}(0) )\\
& \quad \quad ) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{while}.(\ff{and}.( \\
& \quad \mu_0, \ff{eq}.(flag  1).\ff{not}), \\
& \quad \ff{if}.(\mu_1.\ff{not}, \delta_2, \\
& \quad \quad \ff{seq}( \delta_1, flag.\ff{write}(0) )\\
& \quad ) \\
& ) \\
\end{split}
\end{equation}

\begin{theorem}
Expression 7 is semantically equivalent to expression 10:
\end{theorem}

\begin{equation} \nonumber
\begin{split}
& \ff{goto} ( \alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\mu_0, \\
& \quad \quad \ff{if}.(\mu_1, \ff{g}.\ff{forward}(\delta_1), \delta_2) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

\hline

\begin{equation} \nonumber
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{while}.(\ff{and}.( \\
& \quad \mu_0, \ff{eq}.(flag  1).\ff{not}), \\
& \quad \ff{if}.(\mu_1.\ff{not}, \delta_2, \\
& \quad \quad \ff{seq}( \delta_1, flag.\ff{write}(0) \\
& \quad )\\
& ) \\
\end{split}
\end{equation}

To prove that the transformation does not violate program logic, let's consider it with the following abstract case of calling goto forward.

Consider the following code on EO:

\begin{ffcode}
goto
  [g]
    |\sgm{0}| > @ |$\label{ln:ret4}$|
      while.
        |\m{0}|
        |\sgm{1}|
          |\sgm{2}|
          if.
            |\m{1}|
            g.forward |\dlt{1}|
            |\dlt{2}|
          |\sgm{3}||$\label{ln:ret5}$|
\end{ffcode}

@@

Thus, the transformation is:

\begin{ffcode}
memory > flag
flag.write 0
|\sgm{0}||$\label{ln:ret6}$|
  while.
    and.
      (eq. (flag 1)).not
      |\m{0}|
    |\sgm{1}|
      |\sgm{2}|
      if.
        |\m{1}|.not
        |\dlt{2}|
        seq
          |\dlt{1}|
          flag.write 1
      if.
        (eq. (flag 1)).not
        |\sgm{3}|
        TRUE|$\label{ln:ret7}$|
\end{ffcode}

@@

It is worth noting that if goto-forward represent several nested while-loops, then in each of them should be added to the condition of the loop check whether the flag is equal to 1, as in the inner while, from where the jump is made. So, every while that has a nesting less than the original nesting and more than the nesting of the abstract object \emph{g} will end without performing the next iteration.

For example, a program with nested while-loops may look like this:
\begin{ffcode}
goto
  [g]
    while. > @
      |\m{9}|
      |\sgm{9}|
        Y
        |\sgm{4}|
\end{ffcode}

@@

goto
  [g]
    |\sgm{1}| > @
      stdout "Will be printed forever"
      g.backward
\end{ffcode}

Same as with goto forward, there is the goto \emph{atom}-object, the abstract object \emph{g}, and the g.backward attribute. Several abstract objects such as \emph{g} can be declared inside of goto-object, and the attribute \emph{.backward} can be called to each of them.

@@

\begin{ffcode}
goto
  [g]
    g.backward > @
\end{ffcode}

which in terms of $\varphi$-calculus appears as:

\begin{equation}
\begin{split}
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \varphi \mapsto \ff{g}.\ff{backward} \rrbracket ) \\
\end{split}
\end{equation}

This is identical to that discussed in Section 3.1.1. In cases where the jump is performed without condition, you need to add an additional if-statement, similar to the example in Section 3.1.1:

\begin{equation}
\begin{split}
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.( TRUE, \ff{g}.\ff{backward}, TRUE ) \rrbracket) \\
\end{split}
\end{equation}

Same code on EO:

\begin{ffcode}
goto
  [g]
    if. > @
      TRUE
      g.backward
      TRUE
\end{ffcode}

\subsubsection{Goto backward wrapped in an object}
After the conversion described in Section 3.2.1, the simplest call of backward jump will be presented as:

\begin{equation}
\begin{split}
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.( \mu_1, \ff{g}.\ff{backward}, \delta_1 ) \rrbracket) \\
\end{split}
\end{equation}

The transformation under consideration involves the following steps:

\begin{equation}
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\ff{eq}.(flag 0), \\
& \quad \quad \ff{if}.( \mu_1, \ff{g}.\ff{backward}, \delta_1 ) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\ff{eq}.(flag 0), \\
& \quad \quad \ff{seq}(flag.\ff{write}(1), \\
& \quad \quad  \quad \ff{if}.( \mu_1.\ff{not}, \delta_1, flag.\ff{write}(0) ) \\
& \quad \quad ) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{while}.(\ff{eq}.(flag 0), \\
& \quad \ff{seq}(flag.\ff{write}(1), \\
& \quad \quad \ff{if}.( \mu_1.\ff{not}, \delta_1, flag.\ff{write}(0) ) \\
& \quad ) \\
& ) \\
\end{split}
\end{equation}

\begin{theorem}
Expression 13 is semantically equivalent to expression 16:
\end{theorem}

\begin{equation} \nonumber
\begin{split}
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{if}.( \mu_1, \ff{g}.\ff{backward}, \delta_1 ) \rrbracket) \\
\end{split}
\end{equation}

\hline

\begin{equation} \nonumber
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{while}.(\ff{eq}.(flag 0), \\
& \quad \ff{seq}(flag.\ff{write}(1), \\
& \quad \quad \ff{if}.( \mu_1.\ff{not}, \delta_1, flag.\ff{write}(0) ) \\
& \quad ) \\
& ) \\
\end{split}
\end{equation}

To prove that the transformation does not violate program logic, let's consider it with the following abstract case of calling goto backward.

Consider the following code on EO:

@@

Note that goto-backward works the same as do-while construction in other languages: the first iteration of the new loop is always executed, and the next only if the condition is TRUE.
So, all objects that occurs between the declaration of an abstract object \emph{g} and the "jump" itself should be executed before the new while loop once, and all subsequent times were executed depending on the condition of the loop.

Since there is no direct implementation of the do-while construct in EO, but there is an implementation of the while generator, we can resort to a little trick.

To create a do-while emission, we can create a flag globally that will be a condition for our outer loop, whose execution condition will depend on this value. The initial value of this flag will be such that the condition for the execution of the added while loop is fulfilled. Then, in the body of the loop, we change the value of the flag to the opposite. When the jump condition is met, we again change the value of the flag so that as soon as the iteration of the loop is completed, the next iteration will begin, since the loop execution condition will be met.

In order to ensure that when the jump condition is met, all other objects in the body of the newly added while loop are not executed, we wrap them in an additional if-statement, in which we will check whether the jump condition was previously executed.

So, after transformation, the program converts to:
\begin{ffcode}

memory > flag
flag.write 0
while.
  eq. (flag 0)
  seq
    flag.write 1
    |\sgm{1}||$\label{ln:ret10}$|
      |\sgm{2}|
      if.
        (|\m{1}|).not
        |\dlt{1}|
        flag.write 0
      if.
        (eq. (flag 0)).not
        |\sgm{3}|
        TRUE|$\label{ln:ret11}$|

@@

\begin{ffcode}

goto
  [g]
    |\sgm{0}|
      W
      |\sgm{4}|

Applying transformation, we get:
\begin{ffcode}

memory > flag
flag.write 0
while.
  eq. (flag 0)
  seq
    flag.write 1
    |\sgm{0}|
      W'
      if.
        (eq. (flag 0)).not
        |\sgm{4}|
        TRUE

\subsubsection{Goto backward inside the while-loop}
Goto backward also can be called inside a while-loop body.

The simplest way to call goto-backward inside a while-loop is represented as:

\begin{equation}
\begin{split}
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\mu_0, \\
& \quad \quad \ff{if}.(\mu_1, \ff{g}.\ff{backward}, \delta_2) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

The transformation under consideration involves the following steps:

\begin{equation}
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\ff{eq}.(flag 0), \\
& \quad \quad \ff{while}.(\mu_0, \\
& \quad \quad \quad \ff{if}.( \mu_1, \ff{g}.\ff{backward}, \delta_1 ) \\
& \quad \quad ) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\ff{eq}.(flag 0), \\
& \quad \quad \ff{seq}(flag.\ff{write}(1), \\
& \quad \quad \quad \ff{while}.( \\
& \quad \quad \quad \quad \ff{and}.(\mu_0, \ff{eq}.(flag 0).\ff{not}), \\
& \quad \quad \quad \quad \ff{if}.( \mu_1.\ff{not}, \delta_1, flag.\ff{write}(0) ) \\
& \quad \quad \quad ) \\
& \quad \quad ) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

\begin{equation}
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{while}.(\ff{eq}.(flag 0), \\
& \quad \ff{seq}(flag.\ff{write}(1), \\
& \quad \quad \ff{while}.( \\
& \quad \quad \quad \ff{and}.(\mu_0, \ff{eq}.(flag 0).\ff{not}), \\
& \quad \quad \quad \ff{if}.( \mu_1.\ff{not}, \delta_1, flag.\ff{write}(0) ) \\
& \quad \quad ) \\
& \quad ) \\
&  ) \\
\end{split}
\end{equation}

\begin{theorem}
Expression 17 is semantically equivalent to expression 20:
\end{theorem}

\begin{equation} \nonumber
\begin{split}
& \ff{goto}(\alpha_1 \mapsto \llbracket \ff{g} \mapsto \varnothing, \\
& \quad \varphi \mapsto \ff{while}.(\mu_0, \\
& \quad \quad \ff{if}.(\mu_1, \ff{g}.\ff{backward}, \delta_2) \\
& \quad ) \\
& \rrbracket) \\
\end{split}
\end{equation}

\hline

\begin{equation} \nonumber
\begin{split}
& \ff{memory}(flag), \\
& flag.\ff{write}(0), \\
& \ff{while}.(\ff{eq}.(flag 0), \\
& \quad \ff{seq}(flag.\ff{write}(1), \\
& \quad \quad \ff{while}.( \\
& \quad \quad \quad \ff{and}.(\mu_0, \ff{eq}.(flag 0).\ff{not}), \\
& \quad \quad \quad \ff{if}.( \mu_1.\ff{not}, \delta_1, flag.\ff{write}(0) ) \\
& \quad \quad ) \\
& \quad ) \\
&  ) \\
\end{split}
\end{equation}

To prove that the transformation does not violate program logic, let's consider it with the following abstract case of calling goto backward.

As discussed in section 3.1.3, the behavior of while is different from other objects in the language. Therefore, the transformation will be slightly different from that described in section 3.2.2.

To understand that the logic is not broken, let's look at a special case of calling goto backward inside a while-loop.

Let's consider the following code:

\begin{ffcode}
goto
  [g]
    |\sgm{0}| > @|$\label{ln:ret12}$|
      while.
        |\m{0}|
        |\sgm{1}|
          |\sgm{2}|
          if.
            |\m{1}|
            g.backward
            |\dlt{1}|
          |\sgm{3}||$\label{ln:ret13}$|

@@

\begin{ffcode}

memory > flag
flag.write 0
while.
  eq. (flag 0)
  seq
    flag.write 1
    |\sgm{0}||$\label{ln:ret14}$|
      while.
        and.
          (eq. (flag 0)).not
          |\m{0}|
        |\sgm{1}|
          |\sgm{2}|
          if.
            |\m{1}|.not
            |\dlt{1}|
            flag.write 0
          if.
            (eq. (flag 0)).not
            |\sgm{3}|
            TRUE|$\label{ln:ret15}$|

@@

Let's take a look on a program with nested while-loops:
\begin{ffcode}

goto
  [g]
    while. > @
      |\m{9}|
      |\sgm{9}|
        Z
        |\sgm{4}|

After transformation it will look like:
\begin{ffcode}

memory > flag
flag.write 0
while.
  eq. (flag 0)
  seq
    flag.write 1
    while.
      and.
        (eq. (flag 0)).not
        |\m{9}|
      |\sgm{9}|
        Z'
        if.
          (eq. (flag 0)).not
          |\sgm{4}|
          TRUE

\subsection{Final algorithm}
To sum up, we can now fully describe the algorithm itself. The first step is to make the transformations, if necessary, described in sections 3.1.1 and 3.2.1, depending on the current transformation. The second step is going through all the jumps in the program - forward and backward, and depending on the jump we carry out the required transformations: if you need to make a replacement in the while loop, containing a jump - we use transformations 3.1.3 and 3.2.3 accordingly. If you need to make a replacement for a normal object that frames the jump - use transformations 3.1.2 and 3.2.2 respectively.

\section{Conclusion}
As a result of our research, we described an algorithm that automatically converts an input program to an equivalent analog, but without using the \emph{atom} object GOTO. We also showed that the program at the output will contain the same logic as the program at the input.
